---
title: "Data Retrieval from TCGA (LUAD)"
output: html_document
author: "Chaitra Rao"
---

```{r, cache = TRUE, echo=FALSE}
knitr::opts_chunk$set(error = TRUE, cache = TRUE, cache.lazy = FALSE)
```

## Necessary Packages

```{r libararies, message = FALSE}
suppressPackageStartupMessages(library(RTCGAToolbox))
suppressPackageStartupMessages(library(TCGA2STAT))
suppressPackageStartupMessages(library(curatedTCGAData))
suppressPackageStartupMessages(library(devtools))
suppressPackageStartupMessages(library(DT))
```

## Retrieval of data

Retrieval of the following data:
1. RNA Seq gene expression with raw counts
2. DNA Methylation profiles obtained via HumanMethylation450 BeadChip (probes for ca. 450K CpG sites)
3. Somatic non-silent mutations
4. GISTIC-called CNA data summarized by gene -> the quantitative readout can be discreticized -> Ref. to the datatable

```{r LUAD_data, echo=FALSE}
rnaseq.LUAD <- getTCGA(disease = "LUAD", data.type = "RNASeq2", type = "count")
meth.LUAD <- getTCGA(disease = "LUAD", data.type = "Methylation", type = "450K")
mut.LUAD <- getTCGA(disease = "LUAD", data.type = "Mutation", type = "somatic")
cna.LUAD <- curatedTCGAData(diseaseCode = "LUAD",
                        assays = "*GISTIC*",
                        dry.run = FALSE)
cna_LUAD <- getTCGA(disease="LUAD", data.type="CNA_SNP", clinical = TRUE)

```

####Downsize methylation data, include CpG sites that are at the top 1% variance

```{r meth_downsize}
meth.var <- apply(meth.LUAD$dat, 1, var)
meth.data.LUAD <- subset(meth.LUAD$dat, meth.var >= quantile(meth.var, 0.99, na.rm=T) & !is.na(meth.var))

```

####Convert beta to M values

```{r beta2m, echo=FALSE, message=FALSE}
meth.data.LUAD <- beta2m(meth.data.LUAD)
```

####Remove NA values from Methylation data

```{r remove_NA, echo=FALSE}
meth.data.LUAD <- na.omit(meth.data.LUAD)
```

####Print the clinical data in a datamatrix

```{r store_clinical, echo=FALSE}
clin_data.LUAD <- data.matrix(cna.COAD$clinical)
```

#### Getting a datatable for each kind of data

```{r datatable, echo=FALSE}
DT::datatable(head(rnaseq.LUAD$dat))
DT::datatable(head(meth.LUAD$dat))
DT::datatable(head(mut.LUAD$dat))
DT::datatable(head(assays(cna.LUAD)[[2]]))
```

CNA downloaded this way has both normal and tumor readouts; getting rid of normal tissue (that is, getting barcodes with `01` as fourth element at the barcode).

```{r filter_tumor_cna}
is_tumor <- sapply(strsplit(colnames(assays(cna.LUAD)[[2]]), '-'), function(x) return(grepl('^01', x[4])))
cna_tumors <- assays(cna.LUAD)[[2]][,colnames(assays(cna.LUAD)[[2]])[is_tumor]]
```

####Get matched tumor and normal samples

```{r get_matched, echo = FALSE}
rnaseq.LUAD.tum.norm <- TumorNormalMatch(rnaseq.LUAD$dat)
meth.LUAD.tum.norm <- TumorNormalMatch(meth.LUAD$dat)

```

####Print the tumor and normal sample in two separate datatables (dim = gene x sample)

#####RNASeq

```{r print_matched_rnaseq, echo=FALSE}
rnaseq_tum <- (rnaseq.LUAD.tum.norm$primary)
rnaseq_norm <- (rnaseq.LUAD.tum.norm$normal)
```

#####Methylation

```{r print_matched_meth, echo=FALSE}
meth_tum <- (meth.LUAD.tum.norm$primary)
meth_norm <- (meth.LUAD.tum.norm$normal)
```

####Trimming the colnames (TCGA barcodes)

```{r trimming, echo=FALSE, message=FALSE}
string1 <- colnames(rnaseq.LUAD$dat)
colnames(rnaseq.LUAD$dat) <- substring(string1, 1, 12)
string2 <- colnames(meth.data.LUAD)
colnames(meth.data.LUAD) <- substring(string2, 1, 12)
string3 <- colnames(mut.LUAD$dat)
colnames(mut.LUAD$dat) <- substring(string3, 1, 12)
## string4 <- colnames(cna.COAD$dat)
## colnames(cna.COAD$dat) <- substring(string4, 1, 12)
colnames(cna_tumors) <-  substring(colnames(cna_tumors), 1, 12)

```

Ensuring duplicate colnames are not present, and taking them out if so

```{r datamatrices, echo=FALSE}
rnaseq_data <- rnaseq.LUAD$dat
meth_data <- meth.data.LUAD
mut_data <- mut.LUAD$dat
## cna_data <- cna.COAD$dat
##cna_data <- data.matrix(assays(cna.COAD)[[2]])

```

####Making a list out of the data retrieved and getting the coldata

```{r mofalist}
LUAD_data <- list(rnaseq = rnaseq_data,
          meth = meth_data,
          mut = mut_data,
          cna = cna_tumors)
sapply(LUAD_data, function(x) table(duplicated(colnames(x))))
## these are duplicated
sapply(LUAD_data, function(x) colnames(x)[duplicated(colnames(x))])
## not going further on the reasons of the duplications: just getting unique stuff
for (item in names(LUAD_data)) {
    LUAD_data[[item]] <- LUAD_data[[item]][,unique(colnames(LUAD_data[[item]]))]
}
sapply(LUAD_data, function(x) table(duplicated(colnames(x))))
```

```{r columndata}
cdata = as.data.frame(clin_data)
```

####Getting the MOFAobject

```{r mofaobject}
luad_mae <- MultiAssayExperiment(
  experiments = LUAD_data, 
  colData = cdata)

LUAD_data <- createMOFAobject(luad_mae)

```

# Session info

```{r sessionInfo, cache = TRUE}
date()
devtools::session_info()
```




