
```{r, cache = TRUE, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

---
title: "Data Retrieval from TCGA"
output: html_document
author: "Chaitra Rao"
---

## Necessary Packages

```{r libararies, message = FALSE}
suppressPackageStartupMessages(library(RTCGAToolbox))
suppressPackageStartupMessages(library(TCGA2STAT))
suppressPackageStartupMessages(library(curatedTCGAData))
suppressPackageStartupMessages(library(devtools))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lumi))
```

## Retrieval of data

Retrieval of the following data:
1. RNA Seq gene expression with raw counts
2. DNA Methylation profiles obtained via HumanMethylation450 BeadChip (probes for ca. 450K CpG sites)
3. Somatic non-silent mutations
4. GISTIC-called CNA data summarized by gene -> the quantitative readout can be discreticized -> Ref. to the datatable


```{r COAD data, echo=FALSE}

rnaseq.COAD <- getTCGA(disease = "COAD", data.type = "RNASeq2", type = "count", clinical = TRUE)
meth.COAD <- getTCGA(disease = "COAD", data.type = "Methylation", type = "450K", clinical = TRUE)
mut.COAD <- getTCGA(disease = "COAD", data.type = "Mutation", type = "somatic")
cna.COAD <- getTCGA(disease="COAD", data.type="CNA_SNP")
#cna.COAD <- curatedTCGAData(diseaseCode = "COAD",
                        #assays = "*GISTIC*",
                       # dry.run = FALSE)
```



```{r beta2m, echo=FALSE, message=FALSE}
meth.COAD$dat <- beta2m(meth.COAD$dat)
```


Store the clinical data in a datamatrix

```{r store clinical, echo=FALSE}
clin_data <- data.matrix(rnaseq.COAD$clinical)
```


```{r store datatables, echo=FALSE, message=FALSE}
DT::datatable(head(rnaseq.COAD$dat))
DT::datatable(head(meth.COAD$dat))
DT::datatable(head(mut.COAD$dat))
DT::datatable(head(cna.COAD$dat))
```

##Get matched tumor and normal samples

```{r get matched tumor/normal smaples, echo = FALSE}
rnaseq.COAD.tum.norm <- TumorNormalMatch(rnaseq.COAD$dat)
meth.COAD.tum.norm <- TumorNormalMatch(meth.COAD$dat)

```


Store the tumor and normal sample in two separate datatables (dim = gene x sample)

RNASeq
```{r store rnaseq matched, echo=FALSE}
rnaseq_tum <- (rnaseq.COAD.tum.norm$primary)
rnaseq_norm <- (rnaseq.COAD.tum.norm$normal)

```

Methylation
```{r store meth matched, echo=FALSE}
meth_tum <- (meth.COAD.tum.norm$primary)
meth_norm <- (meth.COAD.tum.norm$normal)
```



```{r trimming, echo=FALSE, message=FALSE}
string1 <- colnames(rnaseq.COAD$dat)
colnames(rnaseq.COAD$dat) <- substring(string1, 1, 12)
string2 <- colnames(meth.COAD$dat)
colnames(meth.COAD$dat) <- substring(string2, 1, 12)
string3 <- colnames(mut.COAD$dat)
colnames(mut.COAD$dat) <- substring(string3, 1, 12)
string4 <- colnames(cna.COAD$dat)
colnames(cna.COAD$dat) <- substring(string4, 1, 12)

```



## Getting a datamatrix for each kind of data

Storing the data matrices.

```{r datamatrices, echo=FALSE}
rnaseq_data <- rnaseq.COAD$dat
meth_data <- meth.COAD$dat
mut_data <- mut.COAD$dat
cna_data <- as.matrix(cna.COAD$dat)
#cna_data <- data.matrix(assays(cna.COAD)[[2]])
```

```{r combine data into a list, echo=FALSE, message=FALSE}
COAD_data <- list("rnaseq_data"=rnaseq_data, "meth_data"=meth_data, "mut_data"=mut_data, "cna_data"=cna_data)
```


# Session info

```{r sessionInfo, cache = TRUE}
date()
devtools::session_info()
```




