---
title: "Application of MOFA on BRCA data"
output: html_document
author: "Chaitra Rao"
---

```{r, cache = TRUE, echo=FALSE}
knitr::opts_chunk$set(error = TRUE, cache.lazy = FALSE)
```

## Necessary Packages

```{r libraries, message = FALSE}
suppressPackageStartupMessages(library(RTCGAToolbox))
suppressPackageStartupMessages(library(TCGA2STAT))
suppressPackageStartupMessages(library(curatedTCGAData))
suppressPackageStartupMessages(library(devtools))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lumi))
suppressPackageStartupMessages(library(MOFA))
suppressPackageStartupMessages(library(MultiAssayExperiment))
suppressPackageStartupMessages(library(survival))
```

Saving the genomic information of all the CpG sites.
```{r}
genomic_info <- meth.BRCA$cpgs
```

```{r}
# Loading an existing trained model
MOFAobject <- loadModel("BRCA_model1.hdf5")
MOFAobject
```

## Part 1: Disentangling the heterogeneity: calculation of variance explained by each factor in each view

```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFAobject)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFAobject)
```

```{r}
fout <- plotFactorCor(MOFAobject)
```

## Part 2: Characterisation of individual factors
### Inspection of top weighted features in the active views

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "cna", 
  factors = 1:6,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "mut", 
  factors = 1:6,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "rnaseq", 
  factors = 1:6,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "meth", 
  factors = 1:6,
  show_colnames = FALSE
)
```


If you are only interested in looking at only the top features you can use the `plotTopWeights` function.  
For example, here we plot the mutations with largest loadings on Factor 1. The sign on the right indicates the direction of the loading (positive/negative).
```{r}
plotTopWeights(
  MOFAobject, 
  view="cna", 
  factor=1
)
```


```{r}
plotTopWeights(
  MOFAobject, 
  view="meth", 
  factor=4
)
```

Extracting the genomic information for the important CpGs based on the above plot

```{r}
a <- genomic_info["cg05422029",]
b <- genomic_info["cg02067712",]
c <- genomic_info["cg01016662",]
rbind(a,b,c)
```


```{r}
plotTopWeights(
  MOFAobject, 
  view="mut", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="rnaseq", 
  factor=2
)
```



```{r}
plotWeights(
  MOFAobject, 
  view = "cna", 
  factor = 1, 
  nfeatures = 5
)
```



```{r}
plotWeights(
  MOFAobject, 
  view = "meth", 
  factor = 4, 
  nfeatures = 5
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "mut", 
  factor = 3, 
  nfeatures = 5
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "rnaseq", 
  factor = 2, 
  nfeatures = 5
)
```


```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "cna", 
  factor = 1, 
  features = 5, 
  show_rownames = FALSE
)
```



```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "meth", 
  factor = 4, 
  features = 5, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "mut", 
  factor = 3, 
  features = 5, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "rnaseq", 
  factor = 2, 
  features = 5, 
  show_rownames = FALSE
)
```


# Session info

```{r sessionInfo, cache = TRUE}
date()
devtools::session_info()
```
