---
title: "Application of MOFA on BRCA data"
output: html_document
author: "Chaitra Rao"
---

```{r, cache = TRUE, echo=FALSE}
knitr::opts_chunk$set(error = TRUE, cache.lazy = FALSE)
```

## Necessary Packages

```{r libraries, message = FALSE}
suppressPackageStartupMessages(library(RTCGAToolbox))
suppressPackageStartupMessages(library(TCGA2STAT))
suppressPackageStartupMessages(library(curatedTCGAData))
suppressPackageStartupMessages(library(devtools))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lumi))
suppressPackageStartupMessages(library(MOFA))
suppressPackageStartupMessages(library(MultiAssayExperiment))
suppressPackageStartupMessages(library(PCGSE))
```


```{r}
# Loading an existing trained model
MOFAobject <- loadModel("BRCA_model.hdf5")
MOFAobject
```

## Part 1: Disentangling the heterogeneity: calculation of variance explained by each factor in each view

```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFAobject)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFAobject)
```

## Part 2: Characterisation of individual factors
### Inspection of top weighted features in the active views

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "cna_gain", 
  factors = 1:10,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "cna_loss", 
  factors = 1:10,
  show_colnames = FALSE
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "cna_gain", 
  factor = 5, 
  nfeatures = 15
)
```


If you are only interested in looking at only the top features you can use the `plotTopWeights` function.  
For example, here we plot the mutations with largest loadings on Factor 1. The sign on the right indicates the direction of the loading (positive/negative).
```{r}
plotTopWeights(
  MOFAobject, 
  view="cna_gain", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="cna_loss", 
  factor=5
)
```


```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "cna_gain", 
  factor = 1, 
  features = 20, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "cna_loss", 
  factor = 2, 
  features = 20, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "meth", 
  factor = 3, 
  features = 20, 
  show_rownames = FALSE
)
```

```{r}
plotFactorScatters(
  MOFAobject,
  factors = 1:3
)
```

```{r}
set.seed(1234)
clusters <- clusterSamples(
  MOFAobject, 
  k = 2,        # Number of clusters for the k-means function
  factors = 1   # factors to use for the clustering
)
plotFactorScatter(
  MOFAobject, 
  factors = 1:2, 
  color_by = clusters
)
```


```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns
# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFAobject,
  view = "cna_gain",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```
# Session info

```{r sessionInfo, cache = TRUE}
date()
devtools::session_info()
```
