---
title: "Application of MOFA on BRCA data"
output: html_document
author: "Chaitra Rao"
---

```{r, cache = TRUE, echo=FALSE}
knitr::opts_chunk$set(error = TRUE, cache.lazy = FALSE)
```

## Necessary Packages

```{r libraries, message = FALSE}
suppressPackageStartupMessages(library(RTCGAToolbox))
suppressPackageStartupMessages(library(TCGA2STAT))
suppressPackageStartupMessages(library(curatedTCGAData))
suppressPackageStartupMessages(library(devtools))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(lumi))
suppressPackageStartupMessages(library(MOFA))
suppressPackageStartupMessages(library(MultiAssayExperiment))
suppressPackageStartupMessages(library(survival))
```


```{r}
# Loading an existing trained model
MOFAobject <- loadModel("BRCA_model.hdf5")
MOFAobject
```

## Part 1: Disentangling the heterogeneity: calculation of variance explained by each factor in each view

```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFAobject)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFAobject)
```

```{r}
fout <- plotFactorCor(MOFAobject)
```

## Part 2: Characterisation of individual factors
### Inspection of top weighted features in the active views

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "cna_gain", 
  factors = 1:18,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "cna_loss", 
  factors = 1:18,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "meth", 
  factors = 1:18,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "mut", 
  factors = 1:18,
  show_colnames = FALSE
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "rnaseq", 
  factors = 1:18,
  show_colnames = FALSE
)
```


If you are only interested in looking at only the top features you can use the `plotTopWeights` function.  
For example, here we plot the mutations with largest loadings on Factor 1. The sign on the right indicates the direction of the loading (positive/negative).
```{r}
plotTopWeights(
  MOFAobject, 
  view="cna_gain", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="cna_gain", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="cna_loss", 
  factor=2
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="cna_loss", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="meth", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="mut", 
  factor=8
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="rnaseq", 
  factor=7
)
```


```{r}
plotWeights(
  MOFAobject, 
  view = "cna_gain", 
  factor = 1, 
  nfeatures = 18
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "cna_gain", 
  factor = 4, 
  nfeatures = 18
)
```


```{r}
plotWeights(
  MOFAobject, 
  view = "cna_loss", 
  factor = 2, 
  nfeatures = 18
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "cna_loss", 
  factor = 1, 
  nfeatures = 18
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "meth", 
  factor = 3, 
  nfeatures = 18
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "mut", 
  factor = 8, 
  nfeatures = 18
)
```

```{r}
plotWeights(
  MOFAobject, 
  view = "rnaseq", 
  factor = 7, 
  nfeatures = 18
)
```


```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "cna_gain", 
  factor = 1, 
  features = 18, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "cna_gain", 
  factor = 4, 
  features = 18, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "cna_loss", 
  factor = 2, 
  features = 18, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "cna_loss", 
  factor = 1, 
  features = 18, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "meth", 
  factor = 3, 
  features = 18, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "mut", 
  factor = 8, 
  features = 18, 
  show_rownames = FALSE
)
```

```{r}
plotDataHeatmap(
  MOFAobject, 
  view = "rnaseq", 
  factor = 7, 
  features = 18, 
  show_rownames = FALSE
)
```


```{r}
plotFactorScatters(
  MOFAobject,
  factors = 1:3
)
```

```{r}
set.seed(1234)
clusters <- clusterSamples(
  MOFAobject, 
  k = 2,        # Number of clusters for the k-means function
  factors = 1   # factors to use for the clustering
)
plotFactorScatter(
  MOFAobject, 
  factors = 1:2, 
  color_by = clusters
)
```

```{r survival_analysis}
#Retrieve Survival Data
survivalData = cdata[,3:5]

#replace missing follow up values with days of death
survivalData[is.na(survivalData[,3]),3] = survivalData[is.na(survivalData[,3]),2]

#make a data frame for survival data
survData <- data.frame(Time = as.numeric(survivalData[,3]), Status=as.numeric(survivalData[,1]))

#construct survival Object
SurvObject <-Surv(survData[,1], survData[,2])

fit1 <- survfit(SurvObject ~ cdata$pathologyTstage, data = cdata)
plot(fit1)
```
# Session info

```{r sessionInfo, cache = TRUE}
date()
devtools::session_info()
```
